\simname 2.0 User's Guide

### 2008-10-06 doc/manual.tex
docs and examples
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@740 ce7aba0d-9f81-40a6-a495-cecae84b0515

The latest version of the simulator is available from
\texttt{http://cva.stanford.edu} as a compressed tar archive.  UNIX/Linux
users can extract this archive using the tar utility
\begin{verbatim}
  tar xvfz booksim-1.0.tar.gz  
\end{verbatim}
Windows users can use a compression program such as WinZip to extract
the archive.

The simulator itself is written in C++ and has been specifically
tested with GNU's G++ compiler (version $\ge3$).  In addition, both a
LEX and YACC tool (also known as FLEX and BISON) are needed to create
the configuration parser.  These are standard tools in any UNIX/Linux
development environment.  It is suggested that Windows users download
the CYGWIN versions (\texttt{http://www.cygwin.com}) of these UNIX
development tools to simplify their compilation process.  The
\texttt{Makefile} should be edited so that the first lines give the
paths to the tools.  At Stanford, for example, the compiler, YACC, and
LEX are stored in the \texttt{/usr/pubsw/bin} directory.  The
\texttt{Makefile} reflects this:
\begin{verbatim}
CPP    = /usr/pubsw/bin/g++
YACC   = /usr/pubsw/bin/byacc -d
LEX    = /usr/pubsw/bin/flex
\end{verbatim}
Then, the simulator can be compiled by running \texttt{make} in the
directory that contains the \texttt{Makefile}.

@@

The syntax of the simulator is simply
\begin{verbatim}
  booksim [configfile]
\end{verbatim}
The optional parameter \texttt{configfile} is a file that contains
configuration information for the simulator.  So, for example, to
simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube)
network on uniform traffic, a configuration such as the one shown in
Figure~\ref{fig:config_example} could be used.  This particular
configuration is stored in \texttt{examples/torus88}.

@@

  // Flow control
  num_vcs = 2;

@@

In addition to specifying the topology, the configuration file also
contains basic information about the routing algorithm, flow control,
and traffic.  This simple example uses dimension-order routing and, to
ensure deadlock-freedom of this routing function in the torus, two
virtual channels are required.  The \texttt{injection\_rate} parameter
is added to tell the simulator to inject 0.15 flits per simulation
cycle per node.  Because the simulator operates at the flit level,
most parameters are specified in units of flits as is the case with
the \texttt{injection\_rate}.  Also, any line of the configuration
that begins with \texttt{//} is treated as a comment and ignored by
the simulator.  A detailed list of configuration parameters is given in
Section~\ref{sec:config_params}.

@@

\begin{figure}
\begin{verbatim}
%=================================
% Average latency = 6.02008
% Accepted packets = 0.11 at node 52 (avg = 0.147094)
% latency change    = 1
% throughput change = 1

@@

% Warmed up ...
%=================================
% Average latency = 6.0796
% Accepted packets = 0.119 at node 5 (avg = 0.148266)
% latency change    = 0.00562457
% throughput change = 0.00379387

@@

% Draining all recorded packets ...
% Draining remaining packets ...
====== Traffic class 0 ======
Overall average latency = 6.09083 (1 samples)
Overall average accepted rate = 0.149475 (1 samples)
Overall min accepted rate = 0.138551 (1 samples)
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

After the warm up periods have passed, the simulator prints the
``\texttt{Warmed up}'' message and resets all the simulation statistics.
Then, the measurement phase begins and statistics continue to be
reported after each sample period.  Once the measurement periods have
passed, all the measurement packets are drained from the network
before final latency and throughput numbers are reported.  Details of
the configuration parameters used to control the length of the
simulation phases are covered in Section~\ref{sec:sim_params}.

@@

sim_type       = latency;
injection_rate = 0.1;
\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

Figure~\ref{fig:fly_dist} shows a configuration file that can be used
to determine the distribution of packet latencies in a 2-ary 6-fly
network that uses age-based arbitration.  Note the use of the
\texttt{priority} configuration parameter along with the
\texttt{select} allocators that account for packet priorities.  The
simulator does not output latency distributions by default, but by
editing \texttt{trafficmanager.cpp}, setting the configuration
variable \texttt{DISPLAY\_LAT\_DIST} to true, and recompiling, the
distribution will be displayed at the end of the simulation.  This
technique was used to produced the distribution shown in Figure 25.12
of PPIN.

topology = fly;
k = 2;
n = 6;

routing_function = dest_tag;

vc_allocator = select;
sw_allocator = select;
alloc_iters  = 1;

traffic                = uniform;
const_flits_per_packet = 20;
priority               = age;

sim_type       = latency;
injection_rate = 0.1;
\end{verbatim}
\caption{A configuration file (\texttt{examples/fly26\_age}) for
finding the distribution of packet latencies using age-based
arbitration.}
\label{fig:fly_dist}
\end{figure}

As a final example, Figure~\ref{fig:single} shows the use of the
special single-node topology to test the performance of a switch
allocator --- in this case, the iSLIP allocator.  The
\texttt{in\_ports} and \texttt{out\_ports} options set up a simulation
of an $8\times 8$ crossbar.

topology  = single;
in_ports  = 8;
out_ports = 8;

routing_function = single;

vc_allocator = islip;
sw_allocator = islip;
alloc_iters  = 2;

num_vcs     = 8;
vc_buf_size = 1000;

wait_for_tail_credit = 0;

sim_type       = latency;
injection_rate = 0.1;
\end{verbatim}
\caption{A single-node configuration file (\texttt{examples/single})
for testing the performance of a switch allocator.}
\label{fig:single}
\end{figure}

@@

All information used to configure a simulation is passed through a
configuration file as illustrated by the example in
Section~\ref{sec:run_example}.  This section lists the existing
configuration parameters --- a user can incorporate additional options
by changing the \texttt{booksim\_config.cpp} file.

@@

The \texttt{topology} parameter determines the underlying topology of the
network and the simulator supports four basic topologies:
\begin{opt_list}{single}
\item[fly] A $k$-ary $n$-fly (butterfly) topology. The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter
determines the network's dimension.

@@

\item[torus] A $k$-ary $n$-cube (torus) topology.  The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter determines
the network's dimension.
\end{opt_list}

@@

The \texttt{routing\_function} parameter selects a routing algorithm
for the topology.  Many routing algorithms need multiple virtual
channels for deadlock freedom (VCDF).

\begin{opt_list}{dim\_order\_bal}

\item[dim\_order] Dimension-order routing.  Works for the
\texttt{mesh} topology (1 VCDF) and for the \texttt{torus} topology (2
VCDF).

\item[dim\_order\_bal] Dimension-order routing for the
\texttt{torus} topology with a more balanced use of VCs to
avoid deadlock (2 VCDF).

\item[dim\_order\_ni] A non-interfering version of
dimension-order routing.  Works on the \texttt{torus} or \texttt{mesh}
topology and requires one VC per network terminal.

\item[min\_adapt] A minimal adaptive routing algorithm for
the \texttt{mesh} topology (2 VCDF) and for the \texttt{torus}
topology (3 VCDF).

\item[planar\_adapt] Planar-adaptive routing for the
\texttt{mesh} topology (2 VCDF).  Supports routing around failed channels.

\item[romm] ROMM routing for the \texttt{mesh} (2 VCDF).
Load is balanced by routing in two phases: one from the source to a
random intermediate node in the minimal quadrant and a second from the
intermediate to the destination.

\item[romm\_ni] A non-interfering version of ROMM routing for
the \texttt{mesh} that requires one VC per network terminal.

\item[single] A dummy routing function used for the
\texttt{single} topology.

\item[valiant] Valiant's randomized routing algorithm for the
\texttt{mesh} (2 VCDF) and \texttt{torus} (4 VCDF) topology.

\item[valiant\_ni] A non-interfering version of Valiant's algorithm
for the \texttt{torus} that requires 4 VCs per network terminal.

Also, the simulator code is structured so that additional routing
algorithms can be added with minimal changes to the overall simulator
(see the \texttt{routefunc.cpp} file in the simulator's source code).

@@

\item[vc\_buf\_size] The depth of each virtual in flits.

\item[voq] If non-zero, use virtual-output queuing.  With virtual
output queuing, a separate virtual channel is assigned to each
destination in the network.  This option is most useful when used with
a non-interfering routing algorithm (Section~\ref{sec:routing_algs}).

@@

The rate at which flits are injected into the simulator is set using
the \texttt{injection\_rate} option.  The simulator's cycle time is a
flit cycle, the time it takes a single flit to be injected at a
source, and the injection rate is specified in flits per flit cycle.
For example, setting \texttt{injection\_rate = 0.25} means that each
source injects a new flit one of every four simulator cycles.  The
injection process can also be specified as either Bernoulli
(\texttt{injection\_process = bernoulli}) or an on-off process
(\texttt{injection\_process = on\_off}).  The burstiness of the latter
injection process is controlled via the \texttt{burst\_alpha} and
\texttt{burst\_beta} parameter.  See PPIN Section 24.2.2 for a
description of the on-off process and its parameters.

The unit of injection is packets, which may be comprised of many
flits.  The number of flits per packet is set using the
\texttt{const\_flits\_per\_packet} option.  Each packet may also have an
associated priority, either age-based (\texttt{age}) or none
(\texttt{none}), as specified by the \texttt{priority} option.

The simulator also supports several different traffic patterns that
are specified using the \texttt{traffic} option.  To describe these
patterns, we use the same notation of PPIN Section 3.2: $s_i$ ($d_i$)
denotes the $i^\textrm{th}$ bit of the source (destination) address
whereas $s_x$ ($d_x$) denotes the $x^\textrm{th}$ radix-$k$ digit of
the source (destination) address.  The bit length of an address is $b
= \log_2 N$, where $N$ is the number of nodes in the network.

@@

\item[sample\_period] The sample period is expressed in simulator
cycles and is used as a multiplier when specifying the warm-up length
of a simulation and the maximum number of samples.  Also, intermediate
statistics are displayed once every \texttt{sample\_period} cycles.

\item[warmup\_periods] The length of the simulator warm up expressed
as a multiple of the \texttt{sample\_period}.  After warming up, all
statistics counters are reset.

\item[max\_samples] The total length of simulation expressed as a
multiple of the \texttt{sample\_period}.

@@

\item[reorder] A non-zero value indicates that packet order should be
maintained and reordering time is accounted for in the overall latency.

>>>>

The latest version of the simulator is available from
\texttt{http://cva.stanford.edu} as a compressed tar archive.  UNIX/Linux
users can extract this archive using the tar utility
\begin{verbatim}

  tar xvfz booksim-2.0.tar.gz  
\end{verbatim}

Windows users can use a compression program such as Winrar to extract
the archive.

The simulator itself is written in C++ and has been specifically
tested with GNU's G++ compiler (version $\ge3$).  In addition, both a
LEX and YACC tool (also known as FLEX and BISON) are needed to create
the configuration parser.  These are standard tools in any UNIX/Linux
development environment.  It is suggested that Windows users download
the CYGWIN versions (\texttt{http://www.cygwin.com}) of these UNIX
development tools to simplify their compilation process.  The
\texttt{Makefile} should be edited so that the first lines give the
paths to the tools.  At Stanford, for example, the compiler, YACC, and
LEX are stored in the \texttt{/usr/pubsw/bin} directory.  The
\texttt{Makefile} reflects this:
\begin{verbatim}
CPP    = /usr/pubsw/bin/g++
YACC   = /usr/pubsw/bin/bison -d
LEX    = /usr/pubsw/bin/flex
\end{verbatim}
Then, the simulator can be compiled by running \texttt{make} in the
directory that contains the \texttt{Makefile}.

@@

The syntax of the simulator is simply
\begin{verbatim}
  ./booksim [configfile]
\end{verbatim}
The optional parameter \texttt{configfile} is a file that contains
configuration information for the simulator.  So, for example, to
simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube)
network on uniform traffic, a configuration such as the one shown in
Figure~\ref{fig:config_example} could be used. This particular configuration is stored in examples/torus88.

@@

  // Flow control
  num_vcs = 4;

@@

In addition to specifying the topology, the configuration file also
contains basic information about the routing algorithm, flow control,
and traffic.  This simple example uses dimension-order routing and and four
virtual channels.  The \texttt{injection\_rate} parameter
is added to tell the simulator to inject 0.15 packets per simulation
cycle per node.  The packets are default to size of 1 flit.  Also, any line of the configuration
that begins with \texttt{//} is treated as a comment and ignored by
the simulator.  A detailed list of configuration parameters is given in
Section~\ref{sec:config_params}.

@@

\begin{figure}
\begin{verbatim}
% Average latency = 6.02008
% Accepted packets = 0.11 at node 52 (avg = 0.147094)
% latency change    = 1
% throughput change = 1
>>>>>>> variant B
BEGIN Configuration File
Name: examples/torus88
// Topology
topology = torus;
k = 8;
n = 2;
// Routing
routing_function = dim_order;
// Flow control
num_vcs = 4;
// Traffic
traffic = uniform;
injection_rate = 0.15;

END Configuration File
Unable to open flit watch file, continuing with simulation
0%=================================
% Average latency = 10.6252
% Accepted packets = 0.119 at node 5 (avg = 0.148875)
lat(1) = 10.6252;
bins = [ 0 ...999];
freq = [ 0 0 124 16 497 71 920 166 1336 303 1553 436 1351 
406 926 328 537 213 173 77 43 24 13 6 5 2 1 1 0 ...]

@@

% Warmed up ...Time used is 3000 cycles

@@

% Draining all recorded packets ...
% Draining remaining packets ...
Time taken is 6031 cycles
====== Traffic class 0 ======
Overall average latency = 10.658 (1 samples)
Overall average accepted rate = 0.148383 (1 samples)
Overall min accepted rate = 0.134279 (1 samples)
Average hops = 5.01172 (57148 samples)
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

After the warm up periods have passed (default to 3$\times$sample\_period), the simulator prints the
``\texttt{Warmed up}'' message and resets all the simulation statistics.
Then, the measurement phase begins and statistics continue to be
reported after each sample period.  Once the measurement periods have
passed, all the measurement packets are drained from the network
before final latency and throughput numbers are reported.  Details of
the configuration parameters used to control the length of the
simulation phases are covered in Section~\ref{sec:sim_params}.

@@

sim_type       = latency;
injection_rate = 0.005;
\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

@@

All information used to configure a simulation is passed through a
configuration file as illustrated by the example in
Section~\ref{sec:run_example}.  This section lists the existing
configuration parameters --- a user can incorporate additional options
by changing the \texttt{booksim\_config.cpp} file. Most configuration parameters have default values that should work without user specification

@@

The \texttt{topology} parameter determines the underlying topology of the
network. There is also a set of numerical parameters that describes the size of the networks.

\begin{opt_list}{single}
\item[k] Network radix, the number of routers per dimension
\item[n] Network dimension
\item[c] Network concentration, the number of nodes sharing a single router. Typically set to 1, >1 only in networks that has concentration (i.e. cmesh)
\item[x] (NoC simulations only) The number of routers in the X dimension. Used to calculate channel latency between routers. 
\item[y] (NoC simulations only) The number of routers in the y dimension. Used to calculate channel latency between routers. 
\item[xr] (NoC simulations only) For networks that has c>1, the number of nodes in the x direction per router. Used to calculate channel latency between routers. 
\item[yr] (NoC simulations only) For networks that has c>1, the number of nodes in the y direction per router. Used to calculate channel latency between routers. 
\item[limit] Emulate partially filled networks. Nodes greater or equal to limit does not generate or receive packets during simulation. Though depending on the routing algorithm the routers associated with those nodes may be use to route passing by traffic.

Topologies available in \simname are:
\begin{opt_list}{single}
\item[fly] A $k$-ary $n$-fly (butterfly) topology. The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter
determines the network's dimension.

@@

\item[torus] A $k$-ary $n$-cube (torus) topology.  The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter determines
the network's dimension.
\item[cmesh] Concentrated mesh topology is a A $k$-ary $n$-mesh topology with multiple nodes sharing a single router. The \texttt{c} determines the concentration. The cmesh topology has the option that turns on "express channels" as described in  by default these channels are turned off.

\item[fat tree] Fat Tree topology with 3 levels. Nodes are routers are arranged in a tree structure but the number of links between levels stays constant. At the bottom \texttt{k} nodes shares a level 0 router.

\item[flattened butterfly] A topology based on ***

\item[quad tree] A quad tree topology.

\item[tree 4]

\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung.

\item[MECS] A topology based on (***). Currently it only support 1 flit per packet.

@@

The \texttt{routing\_function} parameter selects a routing algorithm
for the topology.  Many routing algorithms need multiple virtual
channels for deadlock freedom (VCDF). In addition to \texttt{routefunc.cpp}, some topologies source files include additional routing functions. Also, the simulator code is structured so that additional routing algorithms can be added with minimal changes to the overall simulator (see the \texttt{routefunc.cpp} file in the simulator's source code).

@@

\item[vc\_buf\_size] The depth of each virtual channel in flits.

%voq is not currently enabled in the simulator !
%\item[voq] If non-zero, use virtual-output queuing.  With virtual
%output queuing, a separate virtual channel is assigned to each
%destination in the network.  This option is most useful when used with
%a non-interfering routing algorithm (Section~\ref{sec:routing_algs}).

@@

The two traffic modes, injection rate or batch, can be set by the integer paramter \texttt[use\_read\_write], where 0 means injection mode. Injection mode is on by default.

\subsubsection{Injection mode}
The rate at which packkets are injected into the simulator is set using
the \texttt{injection\_rate} option.  The simulator's cycle time is a
flit cycle, the time it takes a single flit to be injected at a
source, and the injection rate is specified in packets per flit cycle.
For example, setting \texttt{injection\_rate = 0.25} means that each
source injects a new packet one of every four simulator cycles.  The
injection process can also be specified as either Bernoulli
(\texttt{injection\_process = bernoulli}) or an on-off process
(\texttt{injection\_process = on\_off}).  The burstiness of the latter
injection process is controlled via the \texttt{burst\_alpha} and
\texttt{burst\_beta} parameter.  See PPIN Section 24.2.2 for a
description of the on-off process and its parameters.

The unit of injection is packets, which may be comprised of many

flits. All packets in injection mode are identical. The number of flits per packet is set using the
\texttt{const\_flits\_per\_packet} option.

\subsubsection[Batch mode]
Batch mode is enabled by setting \texttt{use\_read\_write= 1}. In batch mode, \texttt[injection\_rate] paramter is ignored. Instead, traffic are injected into the network as fast as possible until \texttt{batch\_size} of request packets are generated by each node. When a destination node receives a request packet, it returns a reply packet. The total number of packets injected in batch mode is 2$\times$ \texttt{batch\_size} per node. The parameter \texttt[max\_outstanding] specifies the maximum number of (request or reply) packets waiting at a node to be injected into the network.

\subsubsection[Traffic Patterns]
The simulator also supports several different traffic patterns that
are specified using the \texttt{traffic} option.  To describe these
patterns, we use the same notation of PPIN Section 3.2: $s_i$ ($d_i$)
denotes the $i^\textrm{th}$ bit of the source (destination) address
whereas $s_x$ ($d_x$) denotes the $x^\textrm{th}$ radix-$k$ digit of
the source (destination) address.  The bit length of an address is $b
= \log_2 N$, where $N$ is the number of nodes in the network.

@@

\item[sample\_period] The sample period is expressed in simulator
cycles and is used as a multiplier when specifying the warm-up length
of a simulation and the maximum number of samples.  Also, intermediate
statistics are displayed once every \texttt{sample\_period} cycles. This is only applicable in injection mode.

\item[warmup\_periods] The length of the simulator warm up expressed
as a multiple of the \texttt{sample\_period}.  After warming up, all
statistics counters are reset. This is only applicable in injection mode.

\item[max\_samples] The total length of simulation expressed as a
multiple of the \texttt{sample\_period}. This is only applicable in injection mode.

@@

%This is currently not setup in the traffic manager.
%\item[reorder] A non-zero value indicates that packet order should be
%maintained and reordering time is accounted for in the overall latency.

\item[print\_activity] At the end of a simulation using iq\_router, print out the activity for buffer, switch, and channel of the network.

\item[viewer\_trace] The simulator will generate very verbose print out of all activity inside the network. This print out should be fed into noc\_viewer for a graphic display of the activity inside the network. Currently not working.

\item[watch\_file] Specific flits can have their "watch" status turn on. Require input a file which has flit id listed. 1 id per line.
